---
title: "tidycensus example"
format: pdf
editor: visual
---

## TidyCensus Example

This notebook grabs data from the US Census using the [tidycensus package in R](https://walker-data.com/tidycensus/index.html). Make sure you have installed the `tidyverse` and `tidycensus` packages.

```{r}

library(tidycensus)
library(tidyverse)
```

### Step 1 \| Get a Census API key and replace it in the cell below

```{r}
census_api_key("YOUR-CENSUS-API-KEY-HERE")

```

### Step 2 \| Figure out what tables you want data from

Use <https://censusreporter.org/> to figure out which tables you want.

-   Scroll to the bottom of the page to see the tables.

-   If you already know the table ID, stick that in the "Explore" section to learn more about that table.

Once you've found the variable or variables you want, you can plug them into this code:

```{r}

# the variable B01003_001E was selectd from the census table 
# for population, which we found in censusreporter here:
# https://censusreporter.org/tables/B01003/


# Get variable from ACS
population <- get_acs(geography = "county", 
                      state='NY',
                      variables = c(pop = "B01003_001E"), 
                      year = 2021,
                      survey="acs5")


```

List counties with the highest population in the US

```{r}

population %>% arrange(desc(estimate)) %>% head()
```

### Step 3 \| Example Map

See more details about making maps here:

https://walker-data.com/tidycensus/articles/spatial-data.html

```{r}
options(tigris_use_cache = TRUE)

manhattan <- get_acs(
  state = "NY",
  county = "New York",
  geography = "tract",
  variables = "B01003_001E",
  geometry = TRUE,
  year = 2021
)

manhattan %>%
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c(option = "magma") 


```
